\begin{longtable}{l c c l >{\footnotesize}p{0.28\textwidth}}
	\caption{Literature review of prior nuclei instance segmentation and detection methods.}
	\label{tab:literature-review} \\
	\hline
	\textbf{Method} & \textbf{Year} & \textbf{Author} & \textbf{Approach} & \textbf{Mechanism} \\
	\hline
	\endfirsthead
	\multicolumn{5}{c}{\textit{Table \thetable{} -- continued from previous page}} \\
	\hline
	\textbf{Method} & \textbf{Year} & \textbf{Author} & \textbf{Approach} & \textbf{Mechanism} \\
	\hline
	\endhead
	\hline
	\multicolumn{5}{r}{\textit{Continued on next page}} \\
	\endfoot
	\hline
	\endlastfoot
		StarDist \cite{schmidt2018cell, weigert2022stardist}
			& 2018 & Schmidt et al.
			& contour-based
			& Star-convex polygon; predicts radial distances from centroid to boundary at fixed angular intervals; NMS post-processing \\
		SplineDist \cite{mandal2020splinedist}
			& 2020 & Mandal \& Uhlmann
			& contour-based
			& B-spline control points for cell boundary contours; infers cell shape from radial distances and fits splines; requires centroid-anchored patches; watershed or connected-components post-processing \\
		TSFD-Net \cite{ilyas2022tsfd}
			& 2022 & Ilyas et al.
			& mask-based
			& Tissue-specific feature distillation backbone extracts morphological features varying by tissue type; BiFPN generates hierarchical feature pyramid; interlinked decoders jointly optimize and fuse features; combinational loss for joint nuclei segmentation and classification; NMS post-processing \\
		CPP-Net \cite{chen2023cppnet}
			& 2023 & Chen et al.
			& contour-based
			& Context-aware polygon proposal network; predicts polygon vertices directly with multi-scale feature aggregation; NMS post-processing \\
		CellViT \cite{horst2023cellvit}
			& 2023 & Hörst et al.
			& mask-based
			& Vision Transformer encoder pre-trained on 104M histological image patches; leverages Segment Anything Model foundational architecture; transformer-based segmentation decoder; predicts cell probability, distance maps, and class labels; watershed post-processing \\
		HoVer-NeXt \cite{baumann2024hovernext, graham2019hover}
			& 2024 & Baumann et al.
			& mask-based
			& Horizontal/vertical distance maps from each pixel to neighboring centroids; watershed + morphological post-processing for instance separation \\
		LKCell \cite{cui2024lkcell}
			& 2024 & Cui et al.
			& mask-based
			& Large convolution kernels for computationally efficient receptive fields; transfers pre-trained large-kernel CNN to medical domain; single decoder with NP/HV/NT multitask output heads; watershed post-processing \\
		CellPose \cite{stringer2020cellpose, stringer2025cellposesam}
			& 2020 & Stringer et al.
			& mask-based
			& Gradient flow field pointing toward cell centroids combined with cell probability map; flow-field post-processing \\
		\hline
\end{longtable}
