\chapter{Introduction}

\section{Background}

Tumor-infiltrating lymphocytes (TILs) density has emerged as an established prognostic biomarker in multiple cancer types. The International TILs Working Group has developed standardized guidelines for TILs assessment, which recommend estimating the percentage of stromal area occupied by lymphocytes under a microscope \cite{salgado2015evaluation}. This scoring methodology is fundamentally an \textit{area-based} measurement, which depends on accurately delineating cell boundaries rather than merely locating their centroids.

However, manual TILs assessment presents significant challenges. The process is time-consuming, suffers from substantial inter-observer variability among pathologists, and becomes impractical when applied to whole-slide images (WSIs) at clinical scale. These limitations have motivated the development of automated approaches for nuclei detection and segmentation in histopathology images. 

Critically, practical deployment in clinical settings presents additional constraints: many healthcare facilities, particularly in resource-constrained environments, require models that can run efficiently on edge devices rather than demanding powerful centralized computing infrastructure. This deployment constraint creates a fundamental tension: models must be accurate enough for clinical decision-making yet lightweight enough for practical use in real-world clinical workflows.

Automated approaches for nuclei detection in histopathology images can be broadly categorized into three paradigms, each with distinct trade-offs. \textbf{Instance segmentation methods} such as U-Net \cite{ronneberger2015u} and Mask R-CNN \cite{he2017mask} produce accurate boundary delineations but are computationally expensive, particularly for whole-slide image analysis. \textbf{Vision transformer-based models} like CellViT \cite{cabrielli2023vit} demonstrate strong performance on histopathology tasks, but their self-attention mechanisms impose quadratic complexity in both memory and computational requirements, making them less suitable for resource-constrained deployment. In contrast, \textbf{object detection frameworks} including Faster R-CNN \cite{ren2015faster} and YOLO \cite{redmon2016yolo} offer fast, lightweight inference at the cost of discarding boundary geometry information through bounding box representations, which is insufficient for area-based TILs scoring.

A fundamental challenge across all automated approaches is the \textbf{clumped cell problem}. Densely clustered nuclei are nearly impossible to separate in segmentation masks because fully convolutional network (FCN) based models produce contiguous predictions that merge adjacent cells, leading to systematic undercounting. The computational burden of FCN-based instance segmentation stems not from the convolutional architecture itself, but from the costly post-processing required to separate individual instances. For example, HoVer-Net \cite{graham2019hover} addresses this challenge using horizontal and vertical distance maps, while StarDist \cite{schmidt2018cell} represents cells as star-convex polygons to preserve boundary shape while maintaining per-cell separation. However, these methods still rely on two-stage pipelines (dense prediction followed by post-processing) and typically use non-maximum suppression (NMS), which suppresses overlapping predictions and risks undercounting in precisely the clumped regions where accurate counting matters most.

Recent advances have sought to address the NMS dependency through end-to-end detection frameworks. LSP-DETR \cite{pekar2026lsp} adopts a DETR-style framework with learned assignment mechanisms, eliminating the need for NMS while representing nuclei as star-convex polygons. However, its transformer-based architecture remains computationally demanding for edge deployment scenarios. Meanwhile, CellPose-SAM \cite{stringer2023cellpose} has demonstrated superhuman generalization for cellular segmentation, though its computational profile may not be optimized for real-time edge deployment in clinical workflows.

\textbf{Research Gap:} No existing method combines polygon-based cell representation with a single-stage, end-to-end FCN detection architecture that eliminates NMS dependency while maintaining the lightweight computational profile necessary for edge deployment. This thesis addresses this gap by proposing \textbf{RayCastED}, an FCN-based adaptation of LSP-DETR that re-frames the star-convex representation as a \textit{raycast} representation (centroid plus fixed-angle radial rays), integrated directly into a YOLO-based detection head. By replacing the transformer backbone with an FCN, RayCastED maintains the lightweight profile needed for edge deployment through TensorRT and ONNX Runtime while preserving the end-to-end assigner-based framework that eliminates NMS, thereby addressing both boundary accuracy and clumped-cell separation with detection-level speed.

\subsection*{Key Points Outline}

The following outline summarizes the main points covered in this section:

\begin{enumerate}
	\item \textbf{Clinical significance of TILs.} Tumor-infiltrating lymphocytes (TILs) density is an established prognostic biomarker in several cancers (e.g., the International TILs Working Group guidelines for breast cancer).
	\item \textbf{How TILs are scored.} Pathologists estimate the percentage of stromal area occupied by lymphocytes under a microscope. This is fundamentally an \textit{area-based} measurement, which depends on accurately delineating cell boundaries, not just locating them.
	\item \textbf{Manual problem.} This scoring is time-consuming, suffers from inter-observer variability, and is impractical at whole-slide image (WSI) scale. Furthermore, practical deployment in clinical settings, especially in resource-constrained environments, benefits from models that can run efficiently on edge devices.
	\item \textbf{Automated approaches and trade-offs.}
	\begin{itemize}
		\item Instance segmentation (U-Net, Mask R-CNN): accurate boundaries but computationally expensive
		\item Transformer-based models (CellViT, DINO-based): strong performance but self-attention imposes quadratic complexity in memory and compute. FCN-based architectures will almost always be lighter and more efficient due to their fixed, local receptive fields, making them inherently better suited for resource-constrained edge deployment.
		\item Object detection (YOLO, Faster R-CNN): fast and lightweight but bounding boxes discard boundary geometry, insufficient for area-based scoring.
	\end{itemize}
	\item \textbf{The clumped cell problem.} A fundamental challenge across all approaches: densely clustered nuclei are nearly impossible to separate in segmentation masks because FCN-based models produce contiguous predictions that merge adjacent cells, leading to systematic undercounting. Note that the computational burden of FCN-based instance segmentation stems not from the convolutional architecture itself, but from the costly post-processing required to separate individual instances, which will be discussed in detail in later sections.
	\item \textbf{Polygon-based detection as mitigation.} StarDist and similar methods represent cells as star-convex polygons, preserving boundary shape while maintaining per-cell separation. However, these rely on a two-stage pipeline (dense prediction $\rightarrow$ post-processing) and still use NMS, which suppresses overlapping predictions and risks undercounting in precisely the clumped regions that matter most. LSP-DETR addresses the NMS issue by adopting an end-to-end DETR-style framework with learned assignment, but its transformer-based architecture is computationally demanding for edge deployment.
	\item \textbf{Gap and contribution.} No existing method combines polygon-based cell representation with a single-stage, end-to-end FCN detection architecture that avoids NMS dependency. This thesis proposes \textbf{RayCastED}, an FCN-based adaptation of LSP-DETR that re-frames the star-convex representation as a \textit{raycast} representation (centroid + fixed-angle radial rays), integrated directly into a YOLO-based detection head. By replacing the transformer backbone with an FCN, RayCastED maintains the lightweight profile needed for edge deployment (TensorRT, ONNX Runtime) while preserving the end-to-end assigner-based framework that eliminates NMS, addressing both boundary accuracy and clumped-cell separation with detection-level speed.
\end{enumerate}

\section{Problem Statement}

Based on the background discussed above, the following problems are identified:

\begin{enumerate}
	\item \textbf{Accuracy-deployment trade-off:} Existing methods force a choice between clinical accuracy and practical deployability. Segmentation models provide accurate boundaries but require computational resources unavailable in many clinical settings, while detection models offer lightweight deployment but lack the boundary geometry essential for TILs scoring.
	
	\item \textbf{Clumped cell undercounting:} FCN-based instance segmentation models produce contiguous predictions for densely clustered nuclei, merging adjacent cells and leading to systematic undercounting that compromises both cell counting accuracy and area measurement precision.
	
	\item \textbf{Computational inefficiency:} The computational burden of instance segmentation stems largely from costly post-processing required to separate individual instances, rather than the convolutional architecture itself, making real-time analysis impractical in clinical workflows.
	
	\item \textbf{Loss of boundary geometry:} Object detection models disregard cell boundary geometry by outputting only bounding boxes, which is insufficient for area-based TILs scoring that requires accurate cell delineation for clinical decision-making.
	
	\item \textbf{NMS-based suppression errors:} Methods relying on non-maximum suppression risk undercounting in densely clumped regions by suppressing overlapping predictions, precisely where accurate counting matters most for clinical assessment.
	
	\item \textbf{Deployment inaccessibility:} Vision Transformer-based models achieve better cell separation but impose quadratic computational complexity, making them impractical for resource-constrained edge deployment and limiting accessibility to well-funded institutions only.
\end{enumerate}

\section{Objectives}

To address the identified problems, this thesis has the following objectives:

\begin{enumerate}
	\item Develop \textbf{RayCastED}, a nuclei detection model that:
	\begin{itemize}
		\item preserves cell boundary geometry through a raycast polygon representation (centroid + fixed-angle radial rays), addressing the lack of geometric information in conventional object detection models (Problem 3);
		\item eliminates costly instance separation post-processing by predicting per-cell polygons directly from a detection head, addressing both the clumped cell separation issue (Problem 1) and the post-processing overhead (Problem 2);
		\item replaces NMS with a learned end-to-end assignment mechanism, preventing undercounting in densely clumped regions (Problem 4); and
		\item operates on an FCN backbone to maintain a lightweight computational profile suitable for edge deployment via TensorRT and ONNX Runtime (Problem 5).
	\end{itemize}
	\item Evaluate the performance of RayCastED by comparing it against prior nuclei detection and segmentation models on standard benchmarks.
	\item Evaluate the performance of RayCastED on edge devices.
\end{enumerate}

\section{Research Scope}

This research is bounded by the following scope limitations:

\begin{enumerate}
	\item \textbf{Data characteristics:} The study uses only H\&E stained histopathology images with approximately 0.25 MPP (microns per pixel) resolution from publicly available secondary datasets. The datasets encompass multiple tissue types/organs.
	
	\item \textbf{Computational constraints:} Model evaluation is performed on RTX 5070ti hardware. Baseline comparisons are limited to models compatible with this hardware configuration.
	
	\item \textbf{Validation approach:} The research employs computational validation using standard metrics. Pathologist or expert validation is not included in this study.
	
	\item \textbf{Technical boundaries:} There are no limitations on nuclei types, sizes, or detection scope. No strict constraints are imposed on input patch sizes or memory usage during model training and inference.
\end{enumerate}


%\noindent \textbf{Contoh penulisan batasan skripsi Teknik Elektro:}
%
%%-------------------------------------------------	
%\noindent\fbox{%
%	\parbox{\textwidth}{%
%\begin{enumerate}
%\item Objek penelitian: Studi efisiensi dan kinerja sistem pemantauan suhu dan arus pada sistem pembangkit tenaga listrik.
%\item Metode penelitian: Penelitian eksperimental menggunakan analisis simulasi dan pengujian sistem pada skala laboratorium.
%\item Waktu dan tempat penelitian: Waktu penelitian adalah Maret-Agustus 2022 di 
%FasLab TTL.
%\item Populasi dan sampel: Populasi adalah sistem pembangkit tenaga listrik, dan sampel diambil sebanyak 3 sistem yang berbeda.
%\item Variabel: Variabel bebas adalah metode pemantauan suhu dan arus, dan variabel terikat adalah efisiensi dan kinerja
%\item Hipotesis: bahwa metode pemantauan suhu dan arus berpengaruh terhadap efisiensi dan kinerja sistem pembangkit tenaga listrik.
%\item Keterbatasan Penelitian: Keterbatasan penelitian adalah hanya melibatkan 
%pengujian sistem pada skala laboratorium dan membatasi analisis pada variabel 
%bebas dan terikat.
%\end{enumerate}
%		
%	}%
%}
%
%%-------------------------------------------------	
%\newpage
%
%\noindent \textbf{Contoh penulisan batasan skripsi Teknik Biomedis:}
%
%%-------------------------------------------------	
%\noindent\fbox{%
%	\parbox{\textwidth}{%
%\begin{enumerate}
%\item Objek Penelitian: Studi efektivitas dan keamanan alat elektromedik seperti pacu jantung.
%\item Metode Penelitian: Penelitian deskriptif dan observasional menggunakan survei dan analisis data.
%\item Waktu dan Tempat Penelitian: Waktu penelitian adalah Januari-Juni 2022 di 
%Rumah Sakit Sarjito. 
%\item Populasi dan Sampel: Populasi adalah pasien yang menggunakan pacu jantung, dan sampel diambil dengan metode random sampling sebanyak 100 pasien. 
%\item Variabel: Variabel bebas nya adalah jenis alat pacu jantung, dan variabel terikatnya adalah efektivitas dan keamanan. 
%\item Hipotesis: bahwa jenis alat pacu jantung berpengaruh terhadap efektivitas dan keamanan. 
%\item Keterbatasan Penelitian: Keterbatasan penelitian adalah hanya melibatkan satu rumah sakit sebagai lokasi penelitian dan membatasi dalam analisis data hanya pada variabel bebas dan terikat.
%\end{enumerate}
%		
%	}%
%}
%
%%-------------------------------------------------	
%
%\newpage
%
%\noindent \textbf{Contoh penulisan batasan skripsi Teknologi Informasi:}
%
%%-------------------------------------------------	
%\noindent\fbox{%
%	\parbox{\textwidth}{%
%\begin{enumerate}
%\item Objek Penelitian: Analisis perbandingan efektivitas dan efisiensi antara sistem manajemen proyek tradisional dan sistem manajemen proyek berbasis teknologi informasi. 
%\item Metode Penelitian: Penelitian kualitatif dengan menggunakan wawancara dan 
%survei terhadap para pelaku proyek di berbagai perusahaan. 
%\item Waktu dan Tempat Penelitian: Waktu penelitian adalah Januari-Juni 2022 di 
%perusahaan-perusahaan di wilayah Bantul. 
%\item Populasi dan Sampel: Populasi nya adalah perusahaan yang melakukan proyek, dan sampel diambil sebanyak 10 perusahaan yang menerapkan sistem manajemen 
%proyek tradisional dan 10 perusahaan yang menggunakan sistem manajemen 
%proyek berbasis teknologi informasi. 
%\item Variabel: Variabel bebasnya adalah sistem manajemen proyek, dan variabel 
%terikatnya adalah efektivitas dan efisiensi. 
%\item Hipotesis: bahwa sistem manajemen proyek berbasis teknologi informasi lebih efektif dan efisien dibandingkan dengan sistem manajemen proyek tradisional.
%\item Keterbatasan Penelitian: Keterbatasan penelitian adalah penelitian hanya dilakukan pada perusahaan di wilayah Bantul dan hanya melibatkan wawancara dan survei sebagai metode pengumpulan data.
%\end{enumerate}
%		
%	}%
%}

\section{Research Novelty}

The novelty of this research lies in the development of RayCastED, which represents the first architecture that combines three key elements: (1) star-convex polygon representation for preserving cell boundary geometry, (2) FCN backbone for maintaining lightweight computational profile, and (3) end-to-end learned assignment mechanism that eliminates NMS dependency. This unique combination addresses the fundamental trade-off between boundary accuracy and computational efficiency in nuclei detection for histopathology images.

\section{Research Contributions}

The contributions of this research are as follows:

\begin{enumerate}
	\item \textbf{RayCastED model implementation:} A complete implementation of the proposed polygon-based detection model with raycast representation, integrated into a YOLO-based detection head with end-to-end assignment mechanism.
	
	\item \textbf{Performance benchmarks on standard datasets:} Comprehensive evaluation of RayCastED on publicly available histopathology datasets, comparing performance against baseline models including instance segmentation, object detection, and transformer-based approaches.
	
	\item \textbf{Edge deployment validation:} Validation of RayCastED's deployment capabilities on edge devices through TensorRT and ONNX Runtime conversion, demonstrating practical feasibility for resource-constrained clinical environments.
	
	\item \textbf{Comparison against baseline models:} Systematic comparison with prior nuclei detection and segmentation methods to quantify improvements in accuracy, computational efficiency, and clumped-cell separation performance.
	
	\item \textbf{Reproducible codebase:} A publicly available, well-documented implementation that enables reproducibility and provides a foundation for future research in polygon-based detection for computational pathology.
\end{enumerate}



\section{Presentation Structure}

Sistematika penulisan berisi pembahasan apa yang akan ditulis di setiap bab. 
Sistematika pada umumnya berupa paragraf yang setiap paragraf mencerminkan 
bahasan setiap Bab. Bagian ini akan ditulis setelah seluruh bab dalam skripsi ini selesai disusun.